  
---
title: "Homework 3 solutions"
author: "Wenhao Gou"
output: github_document

---
## 0. Global settings:
```{r setup, message = F, collapse = T}
library(tidyverse)
library(p8105.datasets)
library(lubridate) # For cleaning date object
library(patchwork)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 8, 
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + 
            theme(legend.position = "bottom") +
            theme(plot.title = element_text(hjust = 0.5)))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

```


### 1. Problem 1
#### 1.1 Import the dataset and general overview:
```{r, inport dataset for Q1, collapse = T}
data("instacart")
instacart
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart) ` columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. Deatiled description of the dataset can be found [here](https://p8105.com/dataset_instacart.html)

#### 1.2 Answer of the following questions:

* How many aisles are there, and which aisles are the most items ordered from?

```{r Q1_1, collapse = T}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```

We can see that, there are `134` aisles in total. The aisle which have the largest order number is the fresh vegetables aisle. 

* Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r Q1_2}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, -n)
	) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_bar(stat = "identity", alpha = 0.85, color = "#424242", fill = "#0099B4") +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) +
  ggtitle ( "Number of items ordered in each aisle (>10000)") +
  xlab("")
```


* Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.


```{r Q1_3}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```


* Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)

```{r Q1_4, message = F}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)
```


### 2. Problem 2

This problem uses five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF). The data can be downloaded [here](https://p8105.com/data/accel_data.csv). In this spreadsheet, **variables activity.\* are the activity counts for each minute of a 24-hour day starting at midnight.**

#### 2.1 Load and tidy the dataset

Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).
```{r, inport dataset for Q2, message = F, warning = F}
acc <- read_csv("accel_data.csv") %>%
  janitor::clean_names() %>%
  mutate(day_type = case_when(
    day == "Saturday" ~ "weekend",
    day == "Sunday"   ~ "weekend",
    TRUE              ~ "weekday"
  )) %>%
  pivot_longer(activity_1:activity_1440,
               names_to = "minutes",
               names_prefix = "activity_",
               values_to = "value") %>%
  mutate(minutes = as.numeric(minutes)) %>%
  rename("week_id" = week)
acc
```


After clean and reformat the dataset, the resulting table have `6` variables with `50390` observations. In these 6 variable, the `week_id` indicate the id of the week of the observation, ranging from week 1 to 5. The `day_id` indicate the day number in the record, ranging from 1 to 35 (5 weeks). The variable `day` and `day_type` indicate the information of the day. The variable `minutes` indicate which minutes the  observation was in (in each day, there are 1440 minutes). The final variable `value` indicate the recorded value of the accelerometer. 

#### 2.2 Make a table
Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?


```{r Q2_2.1, message = F}
acc_table <-
  acc %>%
  mutate(day = forcats::
           fct_relevel(day, c("Monday", "Tuesday", "Wednesday", 
                              "Thursday", "Friday", "Saturday", "Sunday"))) %>%
  group_by(week_id, day) %>%
  summarise(total_activity = sum(value))
```

This table is till not that easy for us to analyze. To further tidy this table, we can do the following steps:

```{r Q2_1.2}
acc_table %>% pivot_wider(names_from = week_id,
              values_from = total_activity) %>%
  knitr::kable(align = "r", 
               digits = 0,
               col.names = c("Week No.",1:5))
```

Form the table, we can see that, in general, there are no apparent trend in the data. It is only noticeable that in week no. 4 and 5, on Saturday, the observed individual have a significant low level of activity level compared to other days within that week and between each weeks. 

### 2.3 Make a plot

Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r Q2_3, message = F}
acc %>%
  mutate(day = forcats::
           fct_relevel(day, c("Monday", "Tuesday", "Wednesday", 
                              "Thursday", "Friday", "Saturday", "Sunday")))  %>%
  mutate(hour = (minutes - 1) %/% 60) %>%
  group_by(day, day_id, hour) %>%
  summarise(total_activity_day = mean(value)) %>%
  ggplot(aes(x = hour, y = total_activity_day, group = day_id)) +
  scale_color_brewer(palette = "Paired") +
  #scale_colour_viridis_d(option = "D", end = 0.65, direction = -1) +
  xlim(0, 23) +
  geom_line(aes(color = day), size = 0.75) +
  labs(title = "Daily activity manner") + 
  xlab("Time (h)") + 
  ylab("Average activity level") +
  theme(legend.title = element_blank())
```


From the plot, we can see that, for each day of the week, the observed person was in low activity level form 23:00 to 5:00. After that, the activity level of this people leaped and fluctuated until 20 :00. But it is noticeable that, on Sundays, the activity level of this people is higher than other days in the week from 8:00 to 13:00 and lower after 19:00. Also, on Fridays, the activity level of this people is higher than other days in the week from 18:00 to 22:00.

### 3. Problem 3

#### 3.1 Import and clean the dataset

Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?

```{r Q3_1.1}
# Import and clean the data:

data("ny_noaa")
ny_clean <- ny_noaa %>% 
  mutate(year = year(date),
         month = month(date), 
         day = day(date)) %>%
  mutate(tmax = as.numeric(tmax)/10,
         tmin = as.numeric(tmin)/10) %>%
  select(id:date, year:day, everything())
ny_clean

```

The cleaned dataset have 10 variables with 2595176 observations. The first observation `id` indicate the id of the weather station. The other 4 variables after it indicate the date information, range from 1981.1.1 to 2010.12.31. The variable `prcp` indicate the  precipitation in tenths of mm, `snow` and `snwd` indicated the snowfall and snow depth in mm, `tmax` and `tmin` indicate the maximum and minimum temperature. 

```{r Q3_1.2}
# Statistics of the missing value:
tibble(prcp = sum(is.na(pull(ny_clean, prcp))) / 25951.76,
       snow = sum(is.na(pull(ny_clean, snow))) / 25951.76,
       snwd = sum(is.na(pull(ny_clean, snwd))) / 25951.76,
       tmax = sum(is.na(pull(ny_clean, tmax))) / 25951.76,
       tmin = sum(is.na(pull(ny_clean, tmin))) / 25951.76) %>%
  pivot_longer(prcp:tmin,
               names_to = "Variable" , 
               values_to = "Missing (%)") %>%
  knitr::kable()

```

We can inspect the percentage of missing value for these variables. We can see that, for each of the variable , there are at least 5 percentage of missing value. For the variable `tmin` and `tmax`, there are even greater than 40% percentage of missing value. 

```{r Q3_1.3}
# Most commonly observed value in snowfall (excluded NA)

table(pull(ny_clean,snow)) %>%
  as.data.frame() %>%
  as_tibble() %>%
  arrange(desc(Freq)) %>%
  select(observation = Var1, frequency = Freq)
```

The commonly observed value in snowfall (exclude NA) is 0. This result is reasonable, because only winter will snow. 
 


#### 3.2 Plots

* Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?


```{r Q3_2, warning = F, message = F}
ny_clean %>%
  select(id, year, month, tmax) %>%
  filter(month %in% c(1,7)) %>%
  group_by(id, year, month) %>%
  summarise(aver_tmax = mean(tmax, na.rm = T)) %>%
  ggplot(aes(y = aver_tmax, x = year, color = id)) +
  facet_grid(month~., scales = "free") +
  viridis::scale_color_viridis(discrete = T, option = "A") +
  geom_point(size = 0.5) + 
  geom_path(alpha = 0.5) +
  ggtitle("Average maximum temperature among defferent weather stations cross years") +
  ylab("Average maximum temperature") +
  xlab("Year") +
  theme(legend.position = "none") 
```

We can see that, in general, the error of the average temperature across each station is within 10C. There are just some of outlier in the plot (e.g. 1988 Jun). In general, the mximum tempearture fluctuated each year and stay around -2.5C in Jan and 27.5 in Jun. 


* Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r Q3_3, warning = F, message = F}
p1 <-
  ny_clean %>% 
  select(tmax, tmin) %>%
  pivot_longer(tmax:tmin,
               names_to = "group",
               values_to = "temp") %>%
  ggplot(aes(x = group, y = temp, fill = group)) + 
  geom_violin(alpha = 0.85, size = 0.5) + 
  scale_fill_brewer(palette = "Paired") +
  geom_boxplot(width=0.1, fill = "white") +
  ggtitle("Distribution of max and min temperature") + 
  ylab("Temperature (C)") +
  xlab("") +
  theme(legend.title = element_blank())

p2 <- 
  ny_clean %>% 
  select(year, snow) %>%
  filter(snow > 0) %>%
  filter(snow <100)%>%
  mutate(year = as.factor(year)) %>%
  ggplot(aes(x= snow, y = year)) + 
  geom_density_ridges(alpha = 0.75, color = "#424242", fill = "#0099B4") + 
  ggtitle("Distribution of snowfall (0-100) by year") +
  xlab("Snowfall (mm)") +
  ylab("Year")

p1 + p2
```


We can see that, for the temperature of New York, the median of the maximum temperature is around 20 and minimum temperature is around 0. Also, the distribution of the maximum temperature tend to be left skewed and distribution of the minimum temperate seems to be right skewed. For the distribution of snowfall, we can see that the most common level of snowfall is around 10mm and 28mm. Also, there may be some extreme year that the snowfall level is around 50mm and 78mm. As time goes by, the amount of snowfall tends to decrease as each peak in the plot seems to be shorter and the main peak seemed to be moved leftward. 
